---
title: "Homework 6"
author: Chariz Seijo
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

**Read in the data.**

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

**One city**

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

**Across cities**

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2

**Clean dataframe**

```{r}
baby_df =
  read_csv("./data/birthweight.csv") %>% 
    mutate(
    babysex = as.factor(case_when(
      babysex == 1 ~ "male",
      babysex == 2 ~ "female"
    )),
    malform = as.factor(case_when(
      malform == 0 ~ "absent",
      malform == 1 ~ "present"
    )),
    frace = as.factor(case_when(
      frace == 1 ~ "White",
      frace == 2 ~ "Black",
      frace == 3 ~ "Asian",
      frace == 4 ~ "Puerto Rican",
      frace == 8 ~ "Other",
      frace == 9 ~ "Unknown"
    )),
    mrace = as.factor(case_when(
      mrace == 1 ~ "White",
      mrace == 2 ~ "Black",
      mrace == 3 ~ "Asian",
      mrace == 4 ~ "Puerto Rican",
      mrace == 8 ~ "Other",
      mrace == 9 ~ "Unknown"
    ))
  ) 

sum(is.na(baby_df))
```

**Regression model**

This regression model is built on the assumption that mother's age at delivery and parents' race are associated with birthweight.

```{r}
model_fit = lm(bwt ~ momage + frace + mrace, data = baby_df)

model_fit %>% 
        broom::tidy() %>% 
        knitr::kable()

baby_df %>% 
  add_residuals(model_fit) %>% 
  add_predictions(model_fit) %>% 
  ggplot(aes(x = pred, y = resid)) + 
        geom_point() + 
        geom_smooth(se = FALSE, method = "lm")
```

**Cross-validation**

```{r}
model2 = 
  lm(bwt ~ blength + gaweeks, data = baby_df)

model3 = 
  lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = baby_df)

cv_df =
  crossv_mc(baby_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    model_fit = map(train, ~lm(bwt ~ momage + frace + mrace, data = .x)),
    model2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model3 = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_modelfit = map2_dbl(model_fit, test, ~rmse(model = .x, data = .y)),
    rmse_m2   = map2_dbl(model2, test, ~rmse(model = .x, data = .y)),
    rmse_m3 = map2_dbl(model3, test, ~rmse(model = .x, data = .y)))
```

**RMSE plot**

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```

**Model 3** (head circumference, length, sex, and all interactions) has the best fit to predict baby's birthweight since it has the lowest RMSE.

## Problem 3

**Load weather data**

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

**Generate bootstrap samples**

```{r}
boot_sample = function(df) {
  
  sample_frac(df, replace = TRUE)

}

boot_straps = 
  data_frame(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(weather_df))
  )
```

**Analyzing R squared**

```{r}
rsquared_results =
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 

rsquared_results %>% 
  ggplot(aes(x = r.squared)) + 
  geom_density()
```

This R square plot reveals a normal distribution with a peak at about 0.91.

R square 95% CI: (`r quantile(rsquared_results$r.squared, probs=0.025)`, `r quantile(rsquared_results$r.squared, probs=0.975)`)

**Analyzing log of the product of betas**

```{r}
log_betaf = function(df) {
  
    log(df[1,2]*df[2,2]) %>% 
    tibble() %>% 
    mutate(
      log_betas = .$estimate) %>% 
    select(log_betas)
  
}

log_results =
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy),
    log_betas = map(results, log_betaf)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(log_betas) 

log_results %>% 
  ggplot(aes(x = log_betas)) + 
  geom_density()
```

This plot of the log results also reveals a normal distribution with a peak between 2.00 and 2.03.

Log of the product of betas 95% CI: (`r quantile(log_results$log_betas, probs=0.025)`, `r quantile(log_results$log_betas, probs=0.975)`)